---
title: "Regression Models Course Project"
author: "Daria Kravets"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Assignment Overview

You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

1. “Is an automatic or manual transmission better for MPG?”

2. "Quantify the MPG difference between automatic and manual transmissions".

## Data

Let's have a look at our dataset, specifically on different values of few first rows:

```{r}
library(datasets)
data("mtcars")
head(mtcars)
```

## Automatic vs. manual transmission

First of all, we can look at plot to see if there is any visible difference between MPG values for both types of transmission. Look at appendix 1. 

As we can see, values for different types of transmission are not equally distributed, so that can lead us to conclusion, that we need to explore which one is better.

What can T-test show us?

```{r}
t.test(mtcars$mpg~mtcars$am)
```

P-value is really small, that means we should reject a null-hypothesis, which says that there is no difference between types of transmission. 

In the dataset description it states `Transmission (0 = automatic, 1 = manual)`, so our conclusion is that **manual transmission is better for MPG than automatic**

## Quantify the MPG difference between automatic and manual transmissions

First thing, we need to do a multivariate linear regression:

```{r}
mult <- lm(mpg~., mtcars)
st <- step(mult, direction = "both")
```

And let's have a look at summary

```{r}
summary(st)
```

As we can see, results suggest that best model includes `cyl`, `hp`, `wt` and `am` variables. About 86.59% variance is covered by it. First column (Estimate) shows that number of cylinders reducing MPG value (-3.03 for cyl6 and -2.16 for cyl8), as well as gross horsepower (-0.03) and weight (-2.49). But manual transmission is better than automatic (+1.809).

## Conclusion

In general, manual transmission seems to be better than automatic by 1.809 mpg. But there are some other factors that have impact on mpg values.

## Appendix

#### Appendix 1

```{r}
plot(mtcars$mpg, mtcars$am, pch = 19, col = "#0665a5")
```

#### Appendix 2

```{r}
par(mfrow = c(2, 2))
plot(st)
```
